Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-21 appendix-EPQ.tex
modify

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$. For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$. Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $x$ with priority $p$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{PQ}$. Since $\textit{matched}_{\prec}(u \cdot v,p)$ holds, we can see that $q_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $x$. Then, we can see that $q'_j \xrightarrow{\textit{rm}(x)} q_j$. We already know that that $q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \in \seqPQ$.

@@

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$. For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$. Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $x$ with priority $p$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{PQ}$. Therefore, we can see that $l_2 = u \cdot \textit{put}(x,p) \cdot v \in \seqPQ$.

@@

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(x,p)} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(x)} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into $x \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$. Since (1) $p$ is one of maximal priority in $l_2$, (2) $x$ is the newest item with priority $p$ in $l_2$, and (3) $x$ is not removed until $\textit{rm}(x)$, we know that whether we keep $x$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{PQ}$. Therefore, $l_1 \in \seqPQ$.
\end{itemize}

@@

To prove the $\textit{only if}$ direction, given a data-differentiated execution $e \in \mathcal{I}_{\neq}$. By assumption, it is linearizable with respect to a sequential execution $l \in S$, and the bijection between the operations of $e$ and the method events of $l$ ensures that $l$ is differentiated and belongs to $S_{\neq}$.

@@

\section{Proofs in Section \ref{sec:step-by-step linearizability of extended priority queues}}
\label{sec:appendix in section step-by-step linearizability of extended priority queues}

\subsection{Proof of Lemma \ref{lemma:EPQ is closed under projection}}

>>>>

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$. For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $x \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$. Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only values with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $x$ with priority $p$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{PQ}$. Since $\textit{matched}_{\prec}(u \cdot v,p)$ holds, we can see that $q_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $x$. Then, we can see that $q'_j \xrightarrow{\textit{rm}(x)} q_j$. We already know that that $q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \in \seqPQ$.

@@

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}$. For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $x \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(x,p)} q'_i$. Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only values with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $x$ with priority $p$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{PQ}$. Therefore, we can see that $l_2 = u \cdot \textit{put}(x,p) \cdot v \in \seqPQ$.

@@

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(x,p)} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(x)} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into $x \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$. Since (1) $p$ is one of maximal priority in $l_2$, (2) $x$ is the newest value with priority $p$ in $l_2$, and (3) $x$ is not removed until $\textit{rm}(x)$, we know that whether we keep $x$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{PQ}$. Therefore, $l_1 \in \seqPQ$.
\end{itemize}

@@

To prove the $\textit{only if}$ direction, given a data-differentiated execution $e \in \mathcal{I}_{\neq}$. By assumption, it is linearizable with respect to a sequential execution $l \in S$, and the bijection between the operations of $e$ and the operations of $l$ ensures that $l$ is differentiated and belongs to $S_{\neq}$.

@@

\section{Proofs in Section \ref{sec:checking inclusion by recursive procedure}}
\label{sec:appendix in section checking lnclusion by recursive procedure}

\subsection{Definition of Step-by-Step linearizablity}
\label{sec:appendix definition of step-by-step linearizability}

We introduce the notion of step-by-step linearizability, which means that from a linearization of $e \setminus x$ that satisfy the requirements of priority queue, we can obtain a linearization of $e$ that satisfy the requirements of priority queue. Its formal definition is as follows:

%\vspace{-6pt}
\begin{definition}\label{def:step-by-step linearizability}
Given $\Gamma\in \{\mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, $\Gamma$ is step-by-step linearizability, if for each data-differentiated execution $e$ where $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ for some $\alpha$, then $e \setminus \alpha \sqsubseteq \seqPQ \Rightarrow e \sqsubseteq \seqPQ$.

$\seqPQ$ is step-by-step linearizability, if each $\Gamma$ is step-by-step linearizability. 
\end{definition}

Our notion of step-by-step linearizability is inspired by the step-by-step linearizability in ~\cite{DBLP:conf/icalp/BouajjaniEEH15}.

Given a data-differentiated execution $e$, we can obtain a sequence $e'$ from $e$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$, $\textit{rm}(\textit{empty})$) between each pair of $\textit{call}(\textit{put},a,p)$ and $\textit{ret}(\textit{put},a,p)$ (resp., $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$, $\textit{call}(\textit{rm},\textit{empty})$ and $\textit{ret}(\textit{rm},\textit{empty})$). Such $e'$ is called an execution with linearization points, and we call the projection of $e'$ into $m(a,b)$ the linearization of $e$.

\subsection{Obtaining New Sequences while Ensuring that they are in $\seqPQ$}
\label{sec:obtaining new sequences while ensureing that they are in SeqPQ}

Before we prove the step-by-step linearizability of $\seqPQ$, we introduce several lemmas, which are used to ensure some sub-sequences of $\seqPQ$ still belongs to $\seqPQ$. Given a data-differentiated sequence $l$ and one of its maximal priority $p$, let $O_c(l,p)$ and $O_i(l,p)$ be the set of operations with priorities comparable with $p$ and incomparable with $p$ in $l$, respectively. Similarly we can define $D_c(l,p)$ and $D_i(l,p)$ for se of values instead of set of operations. We can see that each priority of values in $O_i(l,p)$ is either larger or incomparable with priorities of values in  $O_c(l,p)$.

The following lemma shows that if a new sequence is generated by erasing some operations in $O_c(l,p)$ while keeping the remaining $O_c(l,p)$ sub-sequences in $\seqPQ$, then this new sequence is still in $\seqPQ$. Note that this is different from projection on value.

\begin{restatable}{lemma}{EraseOcStillinEPQ}
\label{lemma:erase Oc still in EPQ}
Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l'$ be generated from $l$ by discarding some operations in $O_c(l,p)$, and $l' \vert_{ O_c(l,p) } \in \seqPQ$. Then, $l' \in \seqPQ$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$. Assume that $l'$ is generated from $l$ by discarding $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$. Let $D$ be the set such that $D$ contains $a$, if $\textit{put}(a,\_)$ is in $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$. For each $i$, let $q'_i$ be generated from $q_i$ by erasing values in $D$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{rm}(a) \in O_c(l,p)$: By assumption, $l' \vert_{ O_c(l,p) } \in \seqPQ$. Therefore, $a$ is in $q'_j$ and is the should-be-removed value in $O_c(l,p)$. Since each priority of values in $O_i(l,p)$ is either larger or incomparable with priorities of values in $O_c(l,p)$, we can removed $a_c$ from $q'_j$, and then $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.

\item[-] If $\textit{rm}(a) \in O_i(l,p)$: By assumption we know that $q_j \xrightarrow{\textit{rm}(a)} q_{\textit{j+1}}$. Since $q'_j$ contains the same $D_i(l,p)$ values as $q_j$ and $q'_j$ contains less $D_c(l,p)$ values than $q_j$, we can see that $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
\end{itemize}

For each $q'_{\textit{indj-1}}$, it is easy to see that $q'_{\textit{indj-1}} = q'_{\textit{indj}}$. Therefore, we can see that $l' = o_1 \cdot \ldots o_{\textit{ind1-1}} \cdot o_{\textit{ind1+1}} \cdot \ldots \in \seqPQ$. \qed
\end {proof}

The following lemma shows that if a new sequence is generated by from some time point, erasing operations in $O_i(l,p)$, then this new sequence is still in $\seqPQ$.

\begin{restatable}{lemma}{EraseOiFromSomeTimePointStillinEPQ}
\label{lemma:erase Oi from some time point still in EPQ}
Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l'$ be generated from $l$ by discarding operations in $O_i(l,p)$ from some time point, then, $l' \in \seqPQ$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$. Assume that $l'$ is generated from $l$ by discarding all operations $o_i$ if (1) $o_i \in O_i(l,p)$ and (2) $i \geq k$ for a specific index $k$. Let $D$ be a set such that $a \in D$, if $\textit{put}(a,\_)$ is in $l$ and not in $l'$. For each $0 \leq i \leq m$, let $q'_i$ be generated from $q_i$ by erasing values in $D$.

\item[-] If $o_{f(k)}$ is a $\textit{put}$ operation, then it is obvious that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$. Else, if $o_{f(k)} = \textit{rm}(a)$, we can see $a$ is in $q'_{f(k)-1}$, and since (1) $q'_{f(k)-1}$ contains the same $D_c(l,p)$ values as $q_{f(k)-1}$ and $q'_{f(k)-1}$ contains less $D_i(l,p)$ values than $q_{f(k)-1}$, and (2) each priority of values in $O_i(l,p)$ is either larger or incomparable with priorities of values in  $O_c(l,p)$, we can see that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$. Similarly we can prove the case of $o_{f(j)}$ with $j > k$.
\end{itemize}

\begin{restatable}{lemma}{MakePutWithMaxPriorityHappenEarlier}
\label{lemma:make put with maximal priority happen earlier}
Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l=l_1 \cdot l_2$. Let $l_3$ be the projection of $l_2$ into $\{ \textit{put}(\_,p) \}$, and $l_4$ be the projection of $l_2$ into other operations. Then, $l'=l_1 \cdot l_3 \cdot l_4 \in \seqPQ$.
\end{restatable}

\item[-] For $\textit{n+1} \leq i \leq k$, let $q'_i$ be obtained from $q_n$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ with priority $p$ and in the same order.

\item[-] For $\textit{k+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(i)}$. The order of adding them is the same as $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$.
\end{itemize}

\item[-] For $q'_{\textit{k+1}}$: We already know that $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, and it is easy to see that $q_{f(k+1)-1}$ is obtained from $q_n$ by adding values in $o_{\textit{n+1}} \cdot \ldots \cdot o_{f(k+1)-1}$.

    We can see that $q'_k$ is obtained from $q_n$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, and $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+1)}$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_{f(k+1)}$ is an operation of non-$p$ values, then $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$. Since non-$p$ priority is either smaller or incomparable with $p$, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.

    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some value $a$ with priority $p$. We can see that $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$. Since $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, in $q_{f(k+1)-1}$( and also in $q'_k$), there is no value with priority less than $p$, and $a$ is the first-input value of priority $p$. Therefore, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.
    \end{itemize}

\item[-] For $q'_{\textit{k+i}}$ with $i>1$: We already know that $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, and $q_{f(k+i)-1}$ is obtained from $q_{f(k+i-1)}$ by adding values in $o_{f(k+i-1)+1} \cdot \ldots \cdot o_{f(k+i)-1}$.

    We can see that $q'_{\textit{k+i}-1}$ (resp., $q'_{\textit{k+i}}$) is obtained from $q_{f(k+i-1)}$ (resp., $q_{f(k+i)}$) by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$ (resp., $o_1 \cdot \ldots \cdot o_{f(k+i)}$).

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_{f(k+i)}$ is an operation of non-$p$ values, then $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding values which are (1) with priority $p$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$. Since non-$p$ priority is either smaller or incomparable with $p$, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.

    \item[-] Otherwise, it is only possible that $o_{f(k+i)} = \textit{rm}(a)$ for some value $a$ with priority $p$. We can see that $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding values in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$. Since $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, in $q_{f(k+i)-1}$( and also in $q'_{\textit{k+i-1}}$), there is no value with priority less than $p$, and $a$ is the first-input value of priority $p$. Therefore, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.
    \end{itemize}
\end{itemize}

The following lemma shows that if a new sequence is generated by make some $O_i(l,p)$ behaviors to happen earlier, then this new sequence is still in $\seqPQ$.

\begin{restatable}{lemma}{MakeOiHappenEarlierStillinEPQ}
\label{lemma:make Oi happen earlier still in EPQ}
Given a data-differentiated sequential execution $l \in \seqPQ$ and a maximal priority $p$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l \vert_{ O_i(l,p) } = l_1 \cdot l_2$, let $l' = l_1 \cdot l_3$, where $l_3$ is the projection of $l$ into non-$l_1$ operations. Then, $l' \in \seqPQ$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{PQ}$. Let $l_1 = o'_1 \cdot \ldots \cdot o'_n$, let $l_3 = o'_{\textit{n+1}} \cdot \ldots \cdot o'_m$. Let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$. Let $D$ be the set of values which are added and not removed in $o'_1 \cdot \ldots \cdot o'_n$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] It is easy to see that $l_1 \in \seqPQ$, and let $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_1$ in $\textit{PQ}$.

\item[-] For $\textit{n+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding values in $D$. The order of adding them is the same as $o'_1 \cdot \ldots \cdot o'_n$.
\end{itemize}

\item[-] For $q'_{\textit{n+i}}$: We already know that $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, and it is easy to see that $q_{f(n+i)-1}$ is obtained from $q_{f(n+i-1)}$ by adding $D$-values in $o_{f(n+i-1)+1} \cdot \ldots \cdot q_{f(n+i)-1}$.

     We can see that $q'_{\textit{n+i}-1}$ (resp., $q'_{\textit{n+i}}$) is obtained from $q_{f(n+i-1)}$ (resp., $q_{f(n+i)}$) by adding remanning values in $D$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_{f(n+1)}$ is an operation of $O_c(l,p)$ values, since priority in $O_i(l,p)$ is either larger or incomparable with priority in $O_c(l,p)$, we can see that $q'_n \xrightarrow{o_{f(n+1)}} q'_{\textit{n+1}}$.

    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some value $a$ in $O_c(l,p)$. Since $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, in $q_{f(n+i)-1}$( and also in $q'_{\textit{n+i-1}}$), there is no value with priority less than $p$, and $a$ is the first-input value of priority $p$. Therefore, we can see that $q'_{\textit{n+i-1}} \xrightarrow{o_{f(n+i)}} q'_{\textit{n+i}}$.
    \end{itemize}
\end{itemize}

The following lemma shows that if a new sequence is generated by replacing a prefix with another one which make the priority queue has same content, then this new sequence is still in $\seqPQ$.

\begin{restatable}{lemma}{ReplaceEquivalentPrefixStillinEPQ}
\label{lemma:replace equivalent prefix still in EPQ}
Given a data-differentiated sequential execution $l \in \seqPQ$. Let $l=l_1 \cdot l_2$. Given $l_3 \in \seqPQ$. Assume that the priority queue has same content after executing $l_1$ and $l_3$. Let $l' = l_3 \cdot l_2$. Then, $l' \in \seqPQ$.
\end{restatable}

By assumption we know that $q_k = q'_n$. Then it is not hard to see that $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n \xrightarrow{o_{\textit{k+1}}} q_{\textit{k+1}} \ldots \xrightarrow{o_m} q_m$ is a path in $\textit{PQ}$, and then $l' \in \seqPQ_s$. By Lemma \ref{lemma:EPQ rules and semantics}, we know that $l' \in \seqPQ$. \qed
\end {proof}

\subsection{Proving Step-by-Step Linearizability of $\seqPQ$}
\label{sec:obtaining new sequences while ensureing that they are in SeqPQ}

With the help of Lemma \ref{lemma:erase Oc still in EPQ}, Lemma \ref{lemma:erase Oi from some time point still in EPQ}, Lemma \ref{lemma:make put with maximal priority happen earlier}, Lemma \ref{lemma:make Oi happen earlier still in EPQ} and Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can now prove that $\mathsf{MatchedMaxPriority}$ is step-by-step linearizability. Here we use $\textit{call}(o)$ and $\textit{ret}(o)$ as the call and return action of operation $o$, respectively.

\begin{restatable}{lemma}{EPQ1isStepByStepLinearizability}
\label{lemma:EPQ1 is step-by-step linearizability}
For each data-differentiated execution $e$ where $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ for some $x$, then $e \setminus x \sqsubseteq \seqPQ \Rightarrow e \sqsubseteq \seqPQ$.
\end{restatable}

By assumption we know that there exists $l$, such that $e \sqsubseteq l$ and $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds. Let $p$ be the priority of $x$, then $l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$ for some $u$, $v$ and $w$. Let $e' = e \setminus x$. By assumption there exists sequence $l'$, such that $e' \sqsubseteq l' \in \seqPQ$. Let $e_{\textit{lp}}$ be an execution with linearization points of $e$ and the linearization points is added according to $l'$. Or we can say, $e_{\textit{lp}}$ is generated from $e$ by instrumenting linearization points, and the projection of $e_{\textit{lp}}$ into operations is $l'$. Let $l'_v$ be the shortest prefix of $l'$ that contains all operation of $u \cdot v$.

Let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively. Let us change $O_i(l,p)$ elements in $U$, $V$ and $W$, while keep $O_c(l,p)$ elements unchanged. We proceed by a loop: In the first round of the loop, we start from the first $O_i(l,p)$-value of $l'_v$, and let it be $o_h$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no $O_c(l,p)$-value in $W$ happens before $o_h$. Then, $o_h$ is in the new version of $U \cup V$.

\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $O_c(l,p)$-value $o_w$ in $W$, such that $o_w <_{\textit{hb}} o_h$. Then, in $l'_v$, we put $o_h$ and all $O_i(l,p)$-value whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.

\item[-] Case $3$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is after $\textit{ret}(\textit{rm},x)$. Then, in $l'_v$, we put $o_h$ and all $O_i(l,p)$-value whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
\end{itemize}

In the next round of the loop, we consider the second $O_i(l,p)$-value of $l'_v$, and so on. Our process proceed, until either all element in $l'_v$ are in new version of $U \cup V$, or case $2$ or case $3$ happens and this process terminates. Let $U' \cup V'$ and $W'$ be the new version of $U \cup V$ and $W$ after the process terminates, respectively. Let $O_+$ be the set of operations that are moved into $U' \cup V'$ in the process, and let $O_-$ be the set of operations that are moved into $W'$ in the process.

Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,p)$ while the item is not $x$ in $h$. Let $l''_a$ be the longest prefix of $l'_{\textit{u'v'}}$, where linearization of each operation of $l''_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$. Let $l''_d$ be the projection of $l'$ into operations of $O_x$ that are not in $l''_a$. Let $l''_1 = l''_a \cdot l''_d$. Let $l''_2$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$. Let $l''_3$ be the projection of $l'$ into operations which are not in $l'_{\textit{u'v'}}$. Let $l'' = l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$.

To prove $e \sqsubseteq l''$, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds

\item[-] the operation corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
\end{itemize}

Assume there is a cycle in $G$. According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding operations are in the opposite order in $l''$. Then, we consider all possible case of $o_1$ and $o_2$ as follows: Let $O_a$ and $O_d$ be the set of operations in $l''_a$ and $l''_d$, respectively.

\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,p)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_2 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},x,p)$ in $e_{\textit{lp}}$.
